# Customize R code chunks

#######################################################################################################################
#
# You can customize each R code chunk in your report by providing optional arguments after the r in ```{r},
# which appears at the start of the code chunk. Let's look at one set of options.
#
# R functions sometimes return messages, warnings, and even error messages. By default, R Markdown will
# include these messages in your report. You can use the message, warning and error options to prevent R Markdown
# from displaying these. If any of the options are set to FALSE R Markdown will not include the
# corresponding type of message in the output.
#
# For example, R Markdown would ignore any errors or warnings generated by the chunk below.
#
######################################################################################################################
## Cleaning

For the remainder of the report, we will look only at data from the year 1995. We aggregate our data by location, using the *R* code below.

```{r message = FALSE, warning = FALSE, error = FALSE}

library(nasaweather)
library(dplyr)
library(ggvis)
```

```{r}
year <- 1995

means <- atmos %>%
  filter(year == year) %>%
  group_by(long, lat) %>%
  summarize(temp = mean(temp, na.rm = TRUE),
         pressure = mean(pressure, na.rm = TRUE),
         ozone = mean(ozone, na.rm = TRUE),
         cloudlow = mean(cloudlow, na.rm = TRUE),
         cloudmid = mean(cloudmid, na.rm = TRUE),
         cloudhigh = mean(cloudhigh, na.rm = TRUE)) %>%
  ungroup()
```
